---
title: 'Part 1: Investigation of the exponential distribution in R and comparison   with
  the Central Limit Theorem'
author: "Elena Fedorova"
date: "11 august 2018"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Overview

The key aim of this report is to investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution will be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. For this purpose a thousand of exponential simulations will be performed with lambda = 0.2 for averages of 40 exponentials.

## 2. Simulations

Code simulation with the described parameters:
```{r, echo=TRUE}
set.seed(800)
lambda = 0.2
exponentials = 40
nsim = 1:1000
```

Run the required number of simulations:
```{r, echo=TRUE}
simMeans <- data.frame(x = sapply(nsim, function(x) {mean(rexp(exponentials, lambda))}))
head(simMeans)
```

## 3. Sample Mean versus Theoretical Mean

#Sample Mean
Calculating the mean from the simulations with give the sample mean.
```{r, echo=TRUE}
mean(simMeans$x)
sd(simMeans$x)

```

#Theoretical Mean
The theoretical mean of an exponential distribution is 1/lambda.
```{r, echo=TRUE}
(1/lambda)
```

#Comparison
There is only a slight difference between the simulations sample mean and the exponential distribution theoretical mean.
```{r, echo=TRUE}
abs(mean(simMeans$x)-(1/lambda))

```

## 4. Sample Variance versus Theoretical Variance

#Sample Variance
Calculating the variance from the simulation means with give the sample variance.
```{r, echo=TRUE}
var(simMeans$x)
```

#Theoretical Variance
The theoretical variance of an exponential distribution is (lambda * sqrt(n))^-2.
```{r, echo=TRUE}
(lambda * sqrt(exponentials))^-2
```

#Comparison
There is only a slight difference between the simulations sample variance and the exponential distribution theoretical variance.
```{r, echo=TRUE}
abs(var(simMeans$x)-(lambda * sqrt(exponentials))^-2)
```

## 5. Distribution
Here we'll plot a density histogram for the 1000 simulations. There is an overlay with a normal distribution that has a mean of lambda^-1 and standard deviation of (lambda*sqrt(n))^-1, the theoretical normal distribution for the simulations.

```{r, echo=TRUE}
library(ggplot2)
ggplot(simMeans, aes(x=x)) + 
  geom_histogram(aes(y=..density..), binwidth=0.25, fill="steelblue",color="darkblue") + 
stat_function(fun = dnorm, args = list(mean = (1/lambda), sd=(lambda*sqrt(exponentials))^-1), size = 1, color = "midnightblue") + 
        labs(title="Plot of the Simulations", x="SimMean")
```

